---
title: "Final Project"
output: pdf_document
author: "Collin McMahon"
always_allow_html: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries, include = FALSE}
library(tidyverse)
library(ggplot2)
library(LICORS)
library(mosaic)
library(corrplot)
library(factoextra)
library(arules)
library(arulesViz)
library(gridExtra)
library(knitr)
library(kableExtra)
```

## Abstract

The recommendation systems and the algorithms behind them on streaming services such as Netflix and YouTube are paramount for streamlining the content users see so they do not have to spend immense effort finding what they want. On streaming services such as Netflix, anime is a category of content that is included, as despite being Japanese in origin, anime is quite popular with American audiences. I feel that the recommendation systems are not as effective when it comes to recommending the appropriate anime for me.  In this analysis I will be looking at anime data and the anime watch histories of various individuals to try and find relationships between different anime and their associations with what people watch in order to provide better recommendations. I find this especially relevant because many platforms that solely stream anime lack any sort of recommendation system at all, so it is hard to find content that matches my interests without doing a decent amount of research. To find these relationships, I used both k-means clustering with PCA and association rule mining with the a priori algorithm, to cluster anime into related groups and to find which anime go together based on the lists of what anime individuals have watched. Through the clustering, I discovered two subgroups of anime, the highly mainstream anime and longer anime series, which could potentially be used to give very rough recommendations. If someone mainly watches anime from a given cluster, then recommending them anime from that same cluster would be effective. The association rules proved to be interesting also, yielding many high confidence rules, as well as a decent amount of high lift rules. These high lift rules tended to associate one season of a series with another though, limiting the scope of their use as recommendations, but regardless, many rules were mined that associated different series of anime together. These rules frequently included the most popular anime too, which makes them helpful for recommendation if an individual has not already seen all the popular anime. Overall, this analysis is a starting point at looking at the relationships and associations between different anime, which can be used to construct better performing recommendation systems.

## Introduction

The stay-at-home order has likely shifted the activities people engage in during their free-time. Personally, I started to watch a lot more Netflix as I found myself stuck inside. As someone who watches a mix of American and Japanese content on Netflix, I felt that its recommendation system was missing the mark in terms of recommending anime (Japanese animation) to me. Unlike with American content, where the recommendations appear to be more suited to my specific interests, the anime recommendations tended to simply be what is popular, regardless of the specific preferences I have shown in my history.  
My goal is to gain insight into the relationships between different anime and try to find possible connections or trends in the anime people watch. This information is at the core of building a good recommendation algorithm. I find the recommendations paramount to my experience on these platforms because it helps me easily find content I would enjoy amongst the vast amount of content provided. I would like to use this analysis as a starting point in figuring out what anime are the best to recommend to an individual given their watch history, in order to improve their experience on services that stream anime such as Netflix.

## Methods

The first data set used in this analysis is data about 12,294 anime from [myanimelist.net](myanimelist.net). This data set includes name, genre, type (movie, TV, OVA, etc.), episode count, average rating out of 10, and the number of community members on the site that "follow" each anime. The second data set, also from [myanimelist.net](myanimelist.net), is user preference data from 73,516 users. [myanimelist.net](myanimelist.net) allows each user to add anime they have completed to a list on their profile and give it a rating. This data set includes user ID, anime ID, and rating out of 10 (-1 if no rating was assigned). Entries with the same user ID are all the anime one specific individual has listed as watching on the site.  
\newline
I used two different methods to analyze the relationships between different anime and analyze common patterns in watch histories.  
\newline
The first method is clustering with k-means++ in conjunction with PCA in an attempt to group similar anime together based on the metrics in the first data set such as episode count, average rating, and number of community members. I used PCA here because it would make visualization of the clusters easier, as the clusters would be built from 3 dimensions. Additionally, relatively strong correlations exist in the data set (see Figure 1), especially between rating and community members (see Figure 5 in the Appendix), which makes sense as highly rated anime are expected to be more popular on average. Ideally, PCA would be used on a higher dimension data set, but unfortunately the data on each anime was limited to these three numeric variables. After visual inspection of the clusters, as well as analysis of their centers, I decided 3 clusters best represented this data. These clusters would help summarize this large set of different anime and help group similar anime together beyond simply their genre or type. With this, basic recommendations could be made by recommending anime from the same cluster as the majority of previous anime the user has watched.  

```{r fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "Correlations of Anime Metrics", fig.height = 3, fig.width = 3}
anime_main = read.csv("anime.csv")
rating_main = read.csv("rating.csv")

# Replace anime IDs with their actual name
rating_main$name = anime_main$name[match(rating_main$anime_id, anime_main$anime_id)]
rating_main$name = as.character(rating_main$name)
rating_main = rating_main[c(1, 4, 3)]

# Treat IDs as factors
anime_main$anime_id = as.factor(anime_main$anime_id)
rating_main$user_id = as.factor(rating_main$user_id)

# Treat episodes as numeric
anime_main$episodes = as.numeric(anime_main$episodes)

# summary(anime_main)
# summary(rating_main)

# Remove anime with NA ratings
# colnames(anime_main)[apply(anime_main, 2, anyNA)]
anime_main = na.omit(anime_main)

# qplot(anime_main$rating)

anime_pc = select(anime_main, c(episodes, rating, members))
corr_mat = cor(anime_pc)
corrplot(corr_mat, type = "lower", method = "shade", tl.col = "black", tl.srt = 30,
         tl.cex = 0.5, cl.cex = 0.5)

# qplot(anime_main$rating, anime_main$members)
```

The second method is association rule mining in an attempt to find pairs or groups of anime that were commonly watched together based on the user data. Since each entry of the second data set is a single rating made by a user for an anime, I had to transform the data to group all the anime rated by each user together. After this, I ran the a priori algorithm on the watch lists of every user. Because of the massive amount of potential association rules, I set the a priori algorithm to only consider rules with a support greater than 0.1 and a confidence greater than 0.25. Additionally, I set the maximum amount of anime in a single rule to be 5, in order to keep the number of rules to a manageable amount to analyze.   

## Results

Figure 2 shows the results of using kmeans++ in conjunction with PCA with 3 clusters. Clusters 2 and 3 specifically are quite compact, while cluster 1 is more spread out. Despite this, cluster 1 as well as cluster 3 turn out to represent two important groupings of anime. Table 1 shows the cluster centers represented as z-scores, as well as their sizes. Cluster 1 is the smallest cluster with 467 anime, while cluster 2 is the largest with 7091 anime. Cluster 1 has relatively average episode count, above average ratings, and much above average members. Cluster 2 is below average in all metrics. Cluster 3 has above average episode counts and is slightly above average in ratings. After visual inspection of the cluster scatterplots and analysis of the cluster centers, I decided 3 clusters was the most appropriate for representing this data. Figure 6 in the Appendix shows the results of k-means++ and PCA with 2 and 4 clusters, which provide a lesser visual summary of the data in comparison to 3 clusters.  
\newline
Figure 3 shows the results of a priori on the user preference data set. Out of the 7302 association rules found within the given parameters, many of them appear to have high confidence. The majority of the rules have support between 0.1 and 0.2 and lift between 1 and 5. However, there are a significant amount of rules with lift greater than 5, which will be looked at during the conclusive analysis. Finally, Figure 4 is a graph of these high lift association rules. There is notable clustering in the rules that go together, and incredibly popular anime such as Death Note, Angel Beats, and Sword Art Online can be seen involved in many of these rules.  Figure 7 in the Appendix shows the 20 most popular anime, based on the frequency of their appearance within the user watch histories. Included in the folder with this report are a .graphml file and an interactive HTML document that can be used to visualize this graph in better detail, and look into specific rules with more precision.  

```{r fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "Clusters obtained from k-means++ and then summarized using PCA. The percentages in parentheses represent the variation of the data captured by that principle component"}

# Principle Components Analysis
# z = anime_pc

# pc1 = prcomp(z, scale. = TRUE)

# summary(pc1)
# plot(pc1)

# loadings = pc1$rotation
# scores = pc1$x
# qplot(scores[,1], scores[,2], data = anime_pc,
#       xlab = 'Component 1', ylab = 'Component 2')

# Try to find clusters out of this
x = anime_pc
x = scale(x, center = TRUE, scale = TRUE)

# mu = attr(x, "scaled:center")
# sigma = attr(x, "scaled:scale")

set.seed(3)

# c1 = kmeanspp(x, 2, nstart = 50)

# c1_plot = fviz_cluster(c1, data = x, ellipse.type = "confidence",
#                        geom = c("point"))
# c1_plot + 
#  labs(title = "Plot of Clusters following PCA",
#       x = "Component 1 (47.3%)",
#       y = "Component 2 (32.9%)",
#       caption = "Figure 10: Clusters obtained from k-means++ and then summarized
#       using PCA. The percentages in parentheses represent the variation of the data
#       captured by that principle component.") +
#  theme(plot.title = element_text(hjust = 0.5),
#        plot.caption = element_text(size = 10, hjust = 0.5))
#
c2 = kmeanspp(x, 3, nstart = 50)

c2_plot = fviz_cluster(c2, data = x, ellipse.type = "confidence",
                       geom = c("point"))
c2_plot + 
  labs(title = "Plot of Clusters following PCA",
       x = "Component 1 (47.3%)",
       y = "Component 2 (32.9%)") +
  theme(plot.title = element_text(hjust = 0.5))

# c3 = kmeanspp(x, 4, nstart = 50)

# c3_plot = fviz_cluster(c3, data = x, ellipse.type = "confidence",
#                        geom = c("point"))
# c3_plot +
#   labs(title = "Plot of Clusters following PCA",
#        x = "Component 1 (47.3%)",
#        y = "Component 2 (32.9%)",
#        caption = "Figure 12: Clusters obtained from k-means++ and then summarized
#        using PCA. The percentages in parentheses represent the variation of the data
#        captured by that principle component.") +
#   theme(plot.title = element_text(hjust = 0.5),
#         plot.caption = element_text(size = 10, hjust = 0.5))

# c1$size
# c2$size
# c3$size

# c1$centers
# c2$centers
# c3$centers

```


```{r fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE}
clusters = cbind(c2$centers, c2$size)
rownames(clusters) = c("Cluster 1", "Cluster 2", "Cluster 3")
colnames(clusters)[4] = "size"
knitr::kable(clusters, caption = "Anime Cluster Centers and Sizes")

# anime_main$cluster = c2$cluster

# anime_main %>%
#   group_by(name, rating, genre, cluster) %>%
#   filter(cluster == 2) %>%
#   arrange(desc(rating)) %>%
#   head(10)

# anime_main %>%
#   group_by(name, rating, genre, cluster) %>%
#   filter(cluster == 3) %>%
#   arrange(desc(rating)) %>%
#   head(10)

```

```{r fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "Plot of association rules from the a priori algorithm with min support of 0.1, min confidence of 0.25, and max length of 5"}
### Association Rules

# Barplot of top 20 anime
# watch_counts = rating_main %>%
#   group_by(name) %>%
#   summarize(count = n()) %>%
#   arrange(desc(count))

# head(watch_counts, 20) %>%
#   ggplot() +
#   geom_col(aes(x = reorder(name, count), y = count)) + 
#   coord_flip()

watch_lists = split(x = rating_main$name, f = rating_main$user_id)

watch_lists = lapply(watch_lists, unique)

watch_trans = as(watch_lists, "transactions")
# summary(watch_trans)

anime_rules = apriori(watch_trans, 
                      parameter = list(support = .1, confidence = .25, maxlen = 5),
                      control = list(verbose = FALSE))

# plot(anime_rules)
plot(anime_rules, measure = c("support", "lift"), shading = "confidence",
     main = "Scatterplot of 7302 Anime Association Rules")

# inspect(subset(anime_rules, support > 0.25))
# inspect(subset(anime_rules, lift > 6))

```

```{r fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "Graph of top 100 anime association rules by lift"}
set.seed(7)

sub1 = head(anime_rules, 100, by = 'lift')
plot(sub1, method = 'graph', # engine = "htmlwidget",
     cex = 0.3, arrowSize = 0.15, main = "Graph of 100 Anime Association Rules")

# saveAsGraph(sub1, file = "animerules.graphml")
```

## Conclusions

Both methods of analysis on these data sets provided some interesting insights into the relationships between different anime and the patterns in user watch histories with respect to the anime they watched.  
\newline
Analyzing the results from Table 1, we are able to come up with a rough profile of the types of anime that would be included in each of the clusters. Cluster 1, with above average rating and much above average members, is made up of the highly popular and well-known anime. A potential reason that this cluster is so much smaller than the other two clusters is because highly rated anime are the tail of the ratings distribution, which can be seen in Figure 8. Cluster 2 is below average in all aspects and is the largest cluster, which is unfortunate because using it for recommendations would mean recommending lower rated anime. Cluster 3, with above average episode count and relatively average ratings and members, appears to be made up of the longer anime. Therefore, it is reasonable that this cluster is separated for the other two. Though they are quite large and built on only 3 features, these clusters can be used to generate recommendations by recommending an anime from the same cluster as the majority of anime the user has watched in the past, preferably the highest rated ones. For example, an anime movie that falls in cluster 1 is Koe no Katachi, so if a user has watched a majority of anime belonging to cluster 1, such as Steins:Gate, Clannad, and Kimi no Na Wa, then Koe no Katachi would likely be an appropriate recommendation to make to them. Table 2 shows the top-rated anime in cluster 1 used for the example in this analysis. Additionally, the top-rated anime in cluster 3 are shown in Table 3, though cluster 2 is left out as it is below average in every metric and would not be ideal to use for recommendations. With more features, I feel the precision of the clusters would be increased, which would improve the ability of these clusters to be used for recommendation purposes. However, I feel that the results of the clustering from this analysis provide a good starting point for looking at the differences between anime at a large scale.  
\newline
The results of the association rule mining yielded several thousand rules, many of which were high confidence. High confidence means that there is a high probability of coexistence of the consequent with the antecedent, and these high confidence rules are incredibly important in knowledge discovery. Additionally, there were quite a few rules with lifts above 6 or 7 (see Table 4), which is important because lifts greater than 1 indicate that the antecedent and consequent appear more together than expected. Both high lift and high confidence rules can be used for recommendations because these rules show that the anime they pertain to appear to go well together and appear more often together based on the user preference data. Therefore, if a user has watched the consequent anime in an association rule, recommending them the antecedent anime would be a good idea because those anime appear to go together often. One issue with this analysis was that the highest lift rules tended to be the obvious ones. These rules tended to associate different seasons of the same anime together, which is logical because if a user watched season one, it is highly likely that they watched season 2 if it exists. In the event that they did not watch season 2, then these rules could be used for making good recommendations, but otherwise, the amount of helpful rules is decreased by the rules associating the same series of anime with each other. Another observation that is clear from both the high lift association rules and the graph in Figure 4 is the dominance of the incredibly popular anime in them. This could potentially be because of their high support, but for the users who have already watched a majority of the popular anime, using these rules for recommendations could be less effective. If they have not watched all the popular anime though, these rules will likely serve to be remarkably effective.  
\newline
In conclusion, this analysis helped gain insight into the relationships between different anime and found concrete connections and trends in the anime people watch that could serve as the basis for making recommendations. The heavy associations between different seasons and content of the same anime series were confirmed, but also associations linking unrelated anime were found, which are perfect for using as recommendations. Additionally, a rough clustering of anime into the popular and long groups was made, to further help provide recommendations. I believe that with further research and potentially more factors than the three looked at here, these clusters could be better refined and a better framework for making recommendations could be made. This analysis was a good start and revealed some interesting trends, but with the ever growing amount of content, recommendation algorithms need to continue improving to provide users with content tailored to their interests so they do not have to search in the endless sea of shows and movies.  
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline

## Appendix

```{r fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "Scatterplot of the ratings and number of members of the anime from this data set. The relationship appears to be somewhat exponential, with the majority of anime with large followings having a rating between 7 and 9"}
qplot(anime_main$rating, anime_main$members) +
  labs(title = "Plot of Rating on Members",
       x = "Rating",
       y = "Members") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "Clusters obtained from k-means++ and then summarized using PCA. The percentages in parentheses represent the variation of the data captured by that principle component. These clusters are less interpretable than when k = 3"}
c1 = kmeanspp(x, 2, nstart = 50)

c1_plot = fviz_cluster(c1, data = x, ellipse.type = "confidence",
                       geom = c("point"))
c1_plot = c1_plot +
  labs(title = "Plot of Clusters following PCA",
       x = "Component 1 (47.3%)",
       y = "Component 2 (32.9%)") +
  theme(plot.title = element_text(hjust = 0.5))
  
c3 = kmeanspp(x, 4, nstart = 50)

c3_plot = fviz_cluster(c3, data = x, ellipse.type = "confidence",
                       geom = c("point"))
c3_plot = c3_plot +
  labs(title = "Plot of Clusters following PCA",
       x = "Component 1 (47.3%)",
       y = "Component 2 (32.9%)") +
  theme(plot.title = element_text(hjust = 0.5))
  
grid.arrange(c1_plot, c3_plot, ncol = 2)
```

```{r fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "Names of the 20 most frequent anime found in the lists of anime previously watched by users on myanimelist.net"}
# Barplot of top 20 anime
watch_counts = rating_main %>%
  group_by(name) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

head(watch_counts, 20) %>%
  ggplot() +
  geom_col(aes(x = reorder(name, count), y = count)) +
  labs(title = "Top 20 Anime by Frequency",
       y = "Count",
       x = "Anime Name") +
  coord_flip()
```

```{r fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE}
anime_main$cluster = c2$cluster

cluster1 = anime_main %>%
  group_by(name, rating) %>%
  filter(cluster == 1) %>%
  arrange(desc(rating)) %>%
  head(10)
cluster1 = select(cluster1, c(name, episodes, rating, members))
knitr::kable(cluster1, caption = "Top 10 Cluster 1 Anime by Rating")

cluster3 = anime_main %>%
  group_by(name, rating) %>%
  filter(cluster == 3) %>%
  arrange(desc(rating)) %>%
  head(10)
cluster3 = select(cluster3, c(name, episodes, rating, members))
knitr::kable(cluster3, caption = "Top 10 Cluster 3 Anime by Rating")
```

```{r fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "Histogram of the ratings given to various anime by users on myanimelist.net. Note the peak around 6-6.5 and the major drop on either end."}
qplot(anime_main$rating) +
  labs(title = "Histogram of Rating",
       x = "Rating") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE}
arules = as(anime_rules, "data.frame") %>%
  filter(lift > 6) %>%
  select(-count) %>%
  arrange(desc(lift)) %>%
  mutate_if(is.numeric, ~ round(., 2))
knitr::kable(arules, caption = "Anime Association Rules with Highest Lift") %>% 
  kable_styling(latex_options="scale_down")
```