---
title: "Interactive Association Rule Graph"
author: "Collin McMahon"
output: html_document
always_allow_html: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(LICORS)
library(mosaic)
library(corrplot)
library(factoextra)
library(arules)
library(arulesViz)
```

Hello! This is just a fun little interactive version of my association rule graph that I used to analyze the rules. Using the static one was so disgusting, especially with the overlapping text. This one is super nice because you can zoom in and out, search for specific rules or anime by their title, and see the characteristics of a rule such as their support, confidence, and lifts. Enjoy! :)

```{r message = FALSE, echo = FALSE}
anime_main = read.csv("anime.csv")
rating_main = read.csv("rating.csv")

# Replace anime IDs with their actual name
rating_main$name = anime_main$name[match(rating_main$anime_id, anime_main$anime_id)]
rating_main$name = as.character(rating_main$name)
rating_main = rating_main[c(1, 4, 3)]

# Treat IDs as factors
anime_main$anime_id = as.factor(anime_main$anime_id)
rating_main$user_id = as.factor(rating_main$user_id)

# Treat episodes as numeric
anime_main$episodes = as.numeric(anime_main$episodes)

# Remove anime with NA ratings
anime_main = na.omit(anime_main)

### Association Rules

watch_lists = split(x = rating_main$name, f = rating_main$user_id)

watch_lists = lapply(watch_lists, unique)

watch_trans = as(watch_lists, "transactions")

anime_rules = apriori(watch_trans, 
                      parameter = list(support = .1, confidence = .25, maxlen = 5),
                      control = list(verbose = FALSE))

set.seed(7)

sub1 = head(anime_rules, 100, by = 'lift')
plot(sub1, method = 'graph',  engine = "htmlwidget")
```